---
title: "Regression models - Course weekly tests and textbook excercises"
author: "Maurizio Murino"
date: "6 February 2016"
output: html_document
---


## Test 1

###1.1

Consider the data set given below

```{r1}
x <- c(0.18, -1.54, 0.42, 0.95)
```

And weights given by

```{r2}
w <- c(2, 1, 3, 1)
```

Give the value of μ that minimizes the least squares equation
$$∑n_i=1w_i(x_i−μ)2$$.

__Solution:__

```{r3}
optim(mean(x), function(mu) sum(w * (x - mu)^2), method = "BFGS")$par
```

###1.2

Consider the following data set

``` {r 4}
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
y <- c(1.39, 0.72, 1.55, 0.48, 1.19, -1.59, 1.23, -0.65, 1.49, 0.05)
``` 

Fit the regression through the origin and get the slope treating y
as the outcome and x as the regressor. (Hint, do not center the data since we want regression through the origin, not through the means of the data.)

__Solution:__

```{r 5}
lm(y ~ x -1)
```

###1.3

Do `data(mtcars)` from the `datasets` package and fit the regression model with `mpg` as the outcome and `weight` as the predictor. Give the slope coefficient.

```{r 6}
library(datasets); data(mtcars)
```

__Solution:__

```{r7}
lm(mpg ~ wt, data = mtcars)
```

###1.4

Consider data with an outcome (Y) and a predictor (X). The standard deviation of the predictor is one half that of the outcome. The correlation between the two variables is .5. What value would the slope coefficient for the regression model with Y as the outcome and X as the predictor?

__Solution:__

```{r 8 q1.4}
sy <- 1
sx <- 1/2*sy
cor <- 0.5
cor*sy/sx
```

###1.5

Students were given two hard tests and scores were normalized to have empirical mean 0 and variance 1. The correlation between the scores on the two tests was 0.4. What would be the expected score on Quiz 2 for a student who had a normalized score of 1.5 on Quiz 1?

__Solution:__
```{r q1.5}
1.5*0.4
```

###1.6

Consider the data given by the following

```{r n}
x <- c(8.58, 10.46, 9.01, 9.64, 8.86)
```
What is the value of the first measurement if x were normalized (to have mean 0 and variance 1)?

__Solution:__

```{r 10}
normx <- (x - mean(x)) / sd(x)
normx[1]
```

###1.7

Consider the following data set (used above as well). 

```{r11}
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
y <- c(1.39, 0.72, 1.55, 0.48, 1.19, -1.59, 1.23, -0.65, 1.49, 0.05)
```

What is the intercept for fitting the model with x as the predictor and y as the outcome?

__Solution:__

``` {r q1.7}
summary(lm(y ~ x))
```

###1.8

You know that both the predictor and response have mean 0. What
can be said about the intercept when you fit a linear regression?

```{r q1.8}
x <- rnorm(100)
y <- rnorm(100)
summary(lm(x ~ y))
```

__Solution:__

It must be/close to 0.

###1.9

Consider the data given by

```{r q1.9}
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
```

What value minimizes the sum of the squared distances between these points and itself?

__Solution:__

$$∑n_i=(x_i−μ)^2$$

```{r q1.9b}
optim(mean(x), function(mu) sum((x - mu)^2), method = "BFGS")$par
```



###1.10

Let the slope having fit Y as the outcome and X as the predictor be denoted as β1. Let the slope from fitting X as the outcome and Y as the predictor be denoted as γ1. Suppose that you divide β1 by γ1; in other words consider β1/γ1. What is this ratio always equal to?

__Solution:__

Beta1  = Cor(Y, X) * SD(Y) / SD(X)
Gamma1 = Cor(X, Y) * SD(X) / SD(Y)
Beta1/Gamma1 = Cor(Y, X) / Cor(X, Y) * SD(Y)^2 / SD(X)^2
             = 1 * Var(Y) / Var(X)
             = Var(Y) / Var(X)

-----------------------------------------------------------------

## Test 2

###2.1 

Consider the following data with x as the predictor and y as as the outcome.

```{r 2.1a}
x <- c(0.61, 0.93, 0.83, 0.35, 0.54, 0.16, 0.91, 0.62, 0.62)
y <- c(0.67, 0.84, 0.6, 0.18, 0.85, 0.47, 1.1, 0.65, 0.36)
```

Give a P-value for the two sided hypothesis test of whether β1 from a linear regression model is 0 or not.

__Solution:__
The summary command will outline the coefficient of the linear model that we need.

```{r2.1b}
myfit <- lm(y ~ x)
mycoef <- coef(summary(myfit))
mycoef

# The the pvalue is:
round(mycoef, 3)[8]
```

###2.2
 
Consider the previous problem, give the estimate of the residual standard deviation.

The residual standard deviation is given by $$s_res = √(∑(Y - Y_est)^2/(n - 2))$$

```{r2.2a}
round(summary(myfit)$sigma, 3)
```

###2.3 

In the mtcars data set, fit a linear regression model of weight (predictor) on mpg (outcome). Get a 95% confidence interval for the expected mpg at the average weight. What is the lower endpoint?

__Solution:__

```{r 2.3a}
data(mtcars)
y <- mtcars$mpg
x <- mtcars$wt
myfit2 <- lm(y ~ x, data = mtcars)
round(predict(myfit2,data.frame(x=mean(x)), interval="confidence"), 3)
```

###2.4

Refer to the previous question. Read the help file for mtcars. What is the weight coefficient interpreted as?

__Solution:__

The estimated expected change in mpg per 1,000 lb increase in weight.

###2.5

Consider again the mtcars data set and a linear regression model with mpg as predicted by weight (1,000 lbs). A new car is coming weighing 3000 pounds. Construct a 95% prediction interval for its mpg. What is the upper endpoint?

__Solution:__

```{r 2.5a}
round(predict(myfit2, data.frame(x = 3), interval = ("prediction")), 3)
```

###2.6

Consider again the `mtcars` data set and a linear regression model with `mpg` as predicted by weight (in 1,000 lbs). A “short” ton is defined as 2,000 lbs. Construct a 95% confidence interval for the expected change in mpg per 1 short ton increase in weight. Give the lower endpoint.

__solution:__

```{r 2.6a}
myfit3 <- lm(y ~ I(x/2))
mycoef3 <- coef(summary(myfit3))
round((mycoef3[2,1] + c(-1, 1) * qt(.975, df = myfit3$df) * mycoef3[2, 2]), 3)
```

### 2.7 
If my X from a linear regression is measured in centimeters and I convert it to meters what would happen to the slope coefficient?

__Solution:__

It would remain unaltered.

###2.8

I have an outcome, Y, and a predictor, X and fit a linear regression model with Y=β0+β1X+ϵ to obtain β^0 and β^1. What would be the consequence to the subsequent slope and intercept if I were to refit the model with a new regressor, X+c for some constant, c?

__Solution:__

The new intercept would be  $β^0 − cβ^1$.

###2.9

Refer back to the mtcars data set with mpg as an outcome and weight (wt) as the predictor. About what is the ratio of the the sum of the squared errors, ∑ni=1(Yi−Y^i)2 when comparing a model with just an intercept (denominator) to the model with the intercept and slope (numerator)?

__Solution:__

$Yi^=Y¯Yi^=Y¯$ when the fitted model has an intercept only.

###2.10

Do the residuals always have to sum to 0 in linear regression?

__Solution:__

If an intercept is included: yes.
